The large scale clustering of fast radio bursts in dispersion measure space

### 2015-03-25 writeup/dm_space_distortions.tex
Plotter made

For numerical integration, perform the following change of variables to make
things much easier.
\begin{equation}
C^{dd}_\ell(\chi,\chi')
    =
    \frac{2}{\pi}
    \int_0^{(\chi' + \chi)/2}\ud\bar\chi'
    \int_{\max(-2\bar\chi', -2(\chi -\bar\chi'))}^{\min(2\bar\chi', 2(\chi' -
    \bar\chi'))}\ud\Delta'
    \int_0^\infty\ud k k^2 j_\ell(k(\bar\chi' - \Delta'/2))
    j_{\ell}(k(\bar\chi' + \Delta'/2))
    P_{ee}(k, \bar\chi')
\end{equation}
At this point we could again to the transformation to $C_{\ell\chi\chi'}$
leaving the expression as a triple integral. This is not as computationally
costly as might be expected at first glance, since the $\chi''$ and $\chi'''$
integrals are almost the same for all $\chi$, $\chi'$ pairs with only the
limits of integration changing. Thus the full set of $C_{\ell\chi\chi'}$'s can
be computed with a cumulative sum, only performing one new integral (over $k$)
for each element.

Finally, the cross terms:
\begin{equation}
C^{dx}_\ell(\chi,\chi')
    =
    \frac{2}{\pi}
    \int_0^\chi\ud\chi''
    \int_0^\infty\ud k k^2 
    j_\ell(k\chi') j_{\ell}(k\chi'')P_{ex}(k, (\chi' + \chi'')/2)
\end{equation}
\begin{equation}
    C^{dx}_\ell(\chi,\chi') + C^{xd}_\ell(\chi,\chi')
    =
    \frac{2}{\pi}
    \int_0^\chi\ud\chi''
    \int_0^\infty\ud k k^2 
    j_\ell(k\chi') j_{\ell}(k\chi'')P_{ex}(k, (\chi' + \chi'')/2)
\end{equation}

with the following change of variables for numerical integration:
\begin{align}
C^{dx}_\ell(\chi,\chi')
    =&
    \frac{2}{\pi}
    \int_{-\chi'}^{\chi - \chi'}\ud\Delta'
    \int_0^\infty\ud k k^2 j_\ell(k\chi')
    j_{\ell}(k(\chi'+\Delta'))
    P_{ex}(k, \chi' + \Delta'/2)
    \\
C^{dx}_\ell(\chi,\chi') + C^{xd}_\ell(\chi,\chi')
    \approx&
    \frac{2}{\pi}
    \int_{-\chi'}^{\chi - \chi'}\ud\Delta'
    \int_0^\infty\ud k k^2 j_\ell(k(\chi'-\Delta/2))
    j_{\ell}(k(\chi'+\Delta'/2))
    P_{ex}(k, \chi')
    \nonumber\\
    &+
    \frac{2}{\pi}
    \int_{-\chi}^{\chi' - \chi}\ud\Delta'
    \int_0^\infty\ud k k^2 j_\ell(k(\chi-\Delta/2))
    j_{\ell}(k(\chi+\Delta'/2))
    P_{ex}(k, \chi)
\end{align}
The second line exploits the fact that for $\ell \gg 1$, $\chi \gg \chi'/\ell$ and
$\chi' \gg \chi/\ell$, only $\Delta'\ll\chi$ contributes significantly to the
integral. Also the expressions are very weak functions of $\chi$ and thus
may be shifted by $\Delta/2$ without incurring much error.

>>>>

For numerical integration, perform the following change of variables to make
things much easier.
\begin{align}
C^{dd}_\ell(\chi,\chi')
&=
    \frac{2}{\pi}
    \int_0^{(\chi' + \chi)/2}\ud\bar\chi'
    \int_{\max(-2\bar\chi', -2(\chi -\bar\chi'))}^{\min(2\bar\chi', 2(\chi' -
    \bar\chi'))}\ud\Delta'
    \int_0^\infty\ud k k^2 j_\ell(k(\bar\chi' - \Delta'/2))
    j_{\ell}(k(\bar\chi' + \Delta'/2))
    P_{ee}(k, \bar\chi')
    \\
    &\approx
    \frac{4}{\pi}
    \int_0^{\chi_m}\ud\bar\chi'
    \int_{0}^{\infty}\ud\Delta'
    \int_0^\infty\ud k k^2 j_\ell(k(\bar\chi' - \Delta'/2))
    j_{\ell}(k(\bar\chi' + \Delta'/2))
    P_{ee}(k, \bar\chi')
\end{align}
At this point we could again to the transformation to $C_{\ell\chi\chi'}$
leaving the expression as a triple integral. This is not as computationally
costly as might be expected at first glance, since the $\chi''$ and $\chi'''$
integrals are almost the same for all $\chi$, $\chi'$ pairs with only the
limits of integration changing. Thus the full set of $C_{\ell\chi\chi'}$'s can
be computed with a cumulative sum, only performing one new integral (over $k$)
for each element.

Finally, the cross terms:
\begin{equation}
C^{dx}_\ell(\chi,\chi')
    =
    \frac{2}{\pi}
    \int_0^\chi\ud\chi''
    \int_0^\infty\ud k k^2 
    j_\ell(k\chi') j_{\ell}(k\chi'')P_{ex}(k, (\chi' + \chi'')/2)
\end{equation}

with the following change of variables for numerical integration:
\begin{align}
C^{dx}_\ell(\chi,\chi')
    =&
    \frac{2}{\pi}
    \int_{-\chi'}^{\chi - \chi'}\ud\Delta'
    \int_0^\infty\ud k k^2 j_\ell(k\chi')
    j_{\ell}(k(\chi'+\Delta'))
    P_{ex}(k, \chi' + \Delta'/2)
    \\
C^{dx}_\ell(\chi,\chi') + C^{xd}_\ell(\chi,\chi')
    \approx&
    \frac{2}{\pi}
    \int_{-\chi'}^{\chi - \chi'}\ud\Delta'
    \int_0^\infty\ud k k^2 j_\ell(k(\chi'-\Delta'/2))
    j_{\ell}(k(\chi'+\Delta'/2))
    P_{ex}(k, \chi')
    \nonumber\\
    &+
    \frac{2}{\pi}
    \int_{-\chi}^{\chi' - \chi}\ud\Delta'
    \int_0^\infty\ud k k^2 j_\ell(k(\chi-\Delta'/2))
    j_{\ell}(k(\chi+\Delta'/2))
    P_{ex}(k, \chi)\\
    \approx&
    \frac{4}{\pi}
    \int_{0}^{\infty}\ud\Delta'
    \int_0^\infty\ud k k^2 j_\ell(k(\chi_m-\Delta'/2))
    j_{\ell}(k(\chi_m+\Delta'/2))
    P_{ex}(k, \chi_m)
\end{align}
The second line exploits the fact that for $\ell \gg 1$, $\chi \gg \chi'/\ell$ and
$\chi' \gg \chi/\ell$, only $\Delta'\ll\chi$ contributes significantly to the
integral. Also the expressions are very weak functions of $\chi$ and thus
may be shifted by $\Delta/2$ without incurring much error. The third line
exploits the fact that only one term contributes if $\chi$ and $\chi'$ are
widely separated, and if they are close then the two terms combine to the given
expression.

\red{Actually, with the approximations we've made, we can actually do the
$\Delta$ integrals analytically by transferring all the $\Delta$ to a single
Bessel and shifting the power spectrum to be independent of $\Delta$.  Should
be pretty good for $\ell > 10$.  That is nice because it will control all our
numerical issues with oscillating Bessel functions.}
